---
title: "Chapter 2 Exercises"
author: "Carilli"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    toc: false
    toc_float: false
    df_print: paged
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
```

# {.tabset}

## Exercise 1 {.tabset}

### Data

Read the data

```{r ex1-data}
donuts <- 
here::here("data", "DonutDataX.csv") %>% 
  read_csv()
```

We can use `glimpse()` to get information about each variable

```{r glimpse}

donuts %>% 
  glimpse()
```

### donutsx

We see vector type for each variable.  *donutsx* is a character vector and it should be a numeric vector. Let's `pull` *donutsx*

```{r donutsx-1}
donuts %>% 
  pull(donutsx) 

#  or

donuts$donutsx
```

We can see the 12^th^ observation is "five" and should be 5.  So we will change the observation from "five" to 5 and coerce *donutsx* to a numeric vector.

```{r donutsx-2}
donuts$donutsx[12] <- 5
donuts$donutsx <- as.numeric(donuts$donutsx)
```


### weightx

We can use `DT:datatable` to look at the data.

```{r weightx-1}
donuts %>% 
  DT::datatable()
```


If the problem is not immediately obvious, we could try a visualization like a boxplot

```{r weightx-boxplot}
donuts %>% 
  ggplot(aes(y = weightx)) + 
  geom_boxplot()
```

The error is Patty's weight is listed as 1550 pounds.  That is likely supposed to be 150.  Let's find the position of the wrong weight so we can fix it.

```{r index-weight}
which(donuts$weightx == 1550)
```

Let's fix it.

```{r fix-weight}
donuts$weightx[which(donuts$weightx == 1550)] <- 155

donuts$weightx
```

### childx

We saw in the table above that while Bart and Lisa are children they are coded as -1, that should be 1

```{r}
donuts$childx <- donuts$childx * (-1)

donuts$childx
```

### malex


```{r}
donuts %>% 
  arrange(desc(malex))
```

We see that Marge, Lisa, Patty, and Selma are coded as 1 and the men are coded as 0.  We change this in two ways, rename the variable to *femalex* or subtract *malex* from 1.

```{r malex}
donuts$malex <- 1 - donuts$malex

donuts %>% 
  arrange(desc(malex))
```



## Exercise 2 {.tabset}

### Data 

```{r data}
oly <- 
  here::here("data", "olympics_HW.csv") %>% 
  read_csv() %>% 
  janitor::clean_names()
```

### (a)

```{r ex2-part-a}

oly %>% 
  select(medals, athletes, gdp) %>% 
  summary()

oly %>% 
  select(medals, athletes, gdp) %>% 
  summarytools::descr()
```

### (b)

```{r ex2-part-b}

oly %>% 
  select(country, year, medals, athletes, gdp) %>% 
  slice(n = 1:5)

```

### (c)

```{r ex2-part-c}
n_distinct(oly$year)

oly %>% 
  pull(year) %>% 
  n_distinct()

oly %>% 
  group_by(year) %>% 
  count()
```

### (d)

```{r ex2-part-d}
oly %>% 
  ggplot(aes(x = athletes, y = medals)) + 
  geom_jitter()
```

### (e)

More athletes means more opportunity to win medals. The presence of an athlete is necessary to win a medal.

### (f) 

```{r ex2-part-f}
oly %>% 
  ggplot(aes(x = gdp, y = medals)) +
  geom_jitter()
```

### (g)

```{r ex2-part-g}
oly %>% 
  ggplot(aes(x = population, y = medals)) + 
  geom_jitter()
```

### (h)

```{r ex2-part-h}

oly %>% 
  ggplot(aes(x = temp, y = medals)) + 
  geom_jitter()

```


## Exercise 3 {.tabset}

### Data

```{r data-ex3}
hw <- 
here::here("data", "HeightWage_MenWomenUS_HW.csv") %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  drop_na()
```

### (a)

```{r ex3-part-a}
hw %>% 
  select(wage96, height81, height85, siblings) %>% 
  summarytools::descr() %>% 
  summarytools::tb()
```

### (b)

```{r ex3-part-b}
hw %>% 
  ggplot(aes(x = height85, y = wage96)) +
  geom_jitter() + 
  labs(x = "Adult height (in)", y = "Hourly wages ($)")
```

### (c)

```{r ex3-part-c}
hw %>% 
  filter(wage96 < 500) %>% 
  ggplot(aes(x = height85, y = wage96)) +
  geom_jitter() + 
  labs(x = "Adult height (in)", y = "Hourly wages ($)")
```

### (d)

```{r ex3-part-d}
hw %>% 
  ggplot(aes(x = height81, y = height85)) + 
  geom_jitter() + 
  geom_abline(slope = 1, color ="blue")
```



## Exercise 4 {.tabset}

### Data

```{r ex4-data}
anscome <- 
here::here("data", "AnscombesQuartet.csv") %>% 
  read_csv()
```

### (a)

```{r ex4-part-a}
anscombe %>% 
  summarytools::descr(stats = list("mean", "sd"))

anscombe %>% 
  summarize(across(everything(), list(mean = mean, variance = var))) %>% 
  pivot_longer(everything(),names_to = "measure", values_to = "value")
```

### (b)

```{r ex4-part-b}
anscombe %>% 
  pivot_longer(everything(),
   names_to = c(".value", "set"),
   names_pattern = "(.)(.)"
 ) %>% 
  ggplot(aes(x = x, y = y, color = set)) + 
  geom_point() +
  geom_smooth(method = "lm" , se = FALSE) +
  facet_wrap(~ set) 
```

