---
title: "Exercise 1"
author: "Carilli"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, 
                      warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
```

```{r data}
donuts <- 
here::here("data", "DonutDataX.csv") %>% 
  read_csv()
```
# Answers {.tabset .tabset-fade .tabset-pills}

We can use `glimpse()` to get information about each variable

```{r glimpse}
donuts %>% 
  glimpse()
```

## donutsx

We see vector type for each variable.  *donutsx* is a character vector and it should be a numeric vector. Let's `pull` *donutsx*

```{r donutsx-1}
donuts %>% 
  pull(donutsx) 

#  or

donuts$donutsx
```

We can see the 12^th^ observation is "five" and should be 5.  So we will change the observation from "five" to 5 and coerce *donutsx* to a numeric vector.

```{r donutsx-2}
donuts$donutsx[12] <- 5
donuts$donutsx <- as.numeric(donuts$donutsx)
```


## weightx

We can use `DT:datatable` to look at the data.

```{r weightx-1}
donuts %>% 
  DT::datatable()
```


If the problem is not immediately obvious, we could try a visualization like a boxplot

```{r weightx-boxplot}
donuts %>% 
  ggplot(aes(y = weightx)) + 
  geom_boxplot()
```

The error is Patty's weight is listed as 1550 pounds.  That is likely supposed to be 150.  Let's find the position of the wrong weight so we can fix it.

```{r index-weight}
which(donuts$weightx == 1550)
```

Let's fix it.

```{r fix-weight}
donuts$weightx[which(donuts$weightx == 1550)] <- 155

donuts$weightx
```

## childx

We saw in the table above that while Bart and Lisa are children they are coded as -1, that should be 1

```{r}
donuts$childx <- donuts$childx * (-1)

donuts$childx
```

## malex


```{r}
donuts %>% 
  arrange(desc(malex))
```

We see that Marge, Lisa, Patty, and Selma are coded as 1 and the men are coded as 0.  We change this in two ways, rename the variable to *femalex* or subtract *malex* from 1.

```{r malex}
donuts$malex <- 1 - donuts$malex

donuts %>% 
  arrange(desc(malex))
```


